---
title: "Tips for Teaching Data Science"
subtitel: "UNC Grad Seminar"
resource_files:
- appforthat.jpg
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      highlightLines: yes
      highlightStyle: github
      countIncrementalSlides: false
      ratio: '16:9'

---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
library(tidyverse)
```

```{r xaringan-themer, include=FALSE}
library(xaringanthemer)
style_mono_light(
  base_color = "#26116c",
  text_bold_color = "#fd5e53",
  title_slide_text_color = "#fff8e7",
  background_color = "#fff8e7",
  header_font_google = google_font("Roboto"),
  text_font_google   = google_font("Roboto Condensed"),
  code_font_google   = google_font("Droid Mono")
)
```

class: center, middle, inverse

# 1. Know your audience

---

## Are your students...

--

* ... future Statistics PhD students?

--

* ... future Google Data Scientists?

--

* ... studying something else, and taking a general-purpose DS class?

---

## Decisions that depend on your audience:

* What level of **math** can I assume they know?  

* What **math concepts** will help them in the future?

--

* What level of **statistics** can I assume they know?

* What **statistics concepts** will help them in the future?

--

* What level of **programming** can I assume they know?

* What **programming skills** will help them in the future?

---
class: center, middle, inverse

# 2. Use specific examples, not abstractions

---

## Bad:

Suppose we want to estimate a population mean, $\mu$, from a sample $x_1, ..., x_n$.  

We can make a *confidence interval* for the population mean.

---

## Good:

Suppose you go fishing at Jordan Lake, and catch 30 fish.  You are curious what the true average weight of all the fish in Jordan Lake is.

Let's come up with a *reasonable range* of guesses based on the fish we caught.

---

## Why?

**You** have the advantage of years of specific examples, that you now understand under the "umbrella" of concepts.

--

Your students probably have **no ground to build from** yet.

--

As you go through the course and begin to generalize, you can reference back: 

"This symbol, $\mu$, represents the average of all the fish in Jordan Lake."

---
class: center, middle, inverse

# 3. Leave the math at home - focus on intuition

---

## Bad:

The equation for a **logistic regression** is

$$Y = g^{-1} \left( \beta_0 + \beta X + \epsilon \right)$$

where

$$g = \frac{\log(p)}{\log(1-p)}$$


We estimate $\beta$ by Maximum Likelihood.

---

## Good:

In a **logistic regression**, we assume the residuals are not normal, because the response variable is *categorical*.

We transform our predictions from a linear regression to become *probabilities* of a category.

We choose the coefficients that best predict the observed data.

---
## Why?

* Notation adds to the **cognitive load**.

--

* You and I think the math is beautiful.  Not everyone does.

--

* A Data Scientist doesn't (necessarily) need to understand **how** a coefficient is estimated to be able to responsibly apply a method.

---

## What intuition is important?

* In what settings is this model **appropriate**?

--

* What core **assumptions** does the model rely on?

--

* How do we measure **model success**?

--

* How do we **interpret** model output?



---
class: center, middle, inverse

# 4. Be deliberate about your software choices

---
## Target audience:

* Statistics-focused data science:  R, python, SAS

--

* General (industry) data science: python, R

--

* Data literacy: ???

    - Biology -> R
    - Psychology and Social Science -> SPSS?
    - Full beginners -> Excel/spreadsheets
    
---
## Questions to ponder:

* Is programming a **learning objective** of the course, or a **demonstration tool** for concepts?

--

* What **broader community** will my students belong to?

--

* What language(s) are **you** comfortable with?


---
## Resources:

* RMarkdown or Jupyter

    + *Notebooks:* Combine code chunks and text
    + Teaches reproducibility
    + Crucial teaching tool!
    
--

* RStudio Cloud

    + Skip the package install steps for students
    + Skip the R/RStudio install steps for students
    + Supports jupyter notebooks as well
    + Downside: Only free for small usage
    
--

* Google Colab

    + Like Google Docs, but for jupyter notebooks
    + Free (runs on Google's servers!)
    + Easy to use
    + Downside: Kind of supports R, but not really.
    + Downside: Doesn't teach good file/data management
    
--

* GitHub Classroom

---
class: center, middle, inverse

5. Activities, not lectures

---
## Why?

* There is **actual science** on education techniques.  Most academics ignore it.  It suggests activities > lecture.

--

* **"Learn by Doing"**: Students will develop hands-on skills by practicing.

--

* **"Productive Frustration"**:  By encountering tasks they can't do, students will ask the "right" questions.

--

* It mimics the **real world**.  They are learning this for a reason, not just for intellectual enjoyment!


---
class: center, middle, inverse

# Bonus tip: Protect your own time

---

## Teaching is hard!

* You will **never** be 100% satisfied with your lesson!

--

* Decide how much time you will spend, or what your specific checkpoints are.

--

* You **will** make mistakes.  Use them as teaching moments.

--

* If you go above and beyond, seek **support from your University**:  conference attendance, release time, professional development, teaching awards/grants.

---
class: center, middle

# Don't reinvent the wheel

---
## More resources

* **Data 8** at Berkeley: 

> data8.org

* **R for Data Science** and other free textbooks: 

> r4ds.had.co.nz

* **Data Science in a Box** and other public courses:

> datasciencebox.org

* **Me!**

> github.com/kbodwin

> twitter: @kellybodwin

> www.kelly-bodwin.com

> kelly@bodwin.us
