---
title: "Polish Round Table Network"
subtitle: "Accessing app and data resources."
format: html
editor: visual
theme: minty
---

## Database

The core datasets for this project can be found here:

### [Public Googlesheets Database](https://docs.google.com/spreadsheets/d/1C7Hxx66lVQ433TEFdBLtW_1OSwzjPOBFCvw2FfiwNgA/edit?usp=sharing)

Raw csv files for the dataset can also be downloaded from here:

### [Data csv file page](https://zenodo.org/record/7837475#.ZD1OouxBzIg)

#### Submit a correction or addition

Did you find a mistake in the data? Or do you have new information that can be added to enrich the dataset?

If so, please submit your information, and a thorough documentation of your source(s), at this form:

### [Dataset Improvement Form](https://forms.gle/tFUTZxUTArkhbFiZ6)

## Shiny Application for Visualization

To use the `Shiny` app to visualize networks over time, you must first install our custom R package, `PolandRoundTable`. You will need a recent version of R (ideally, 4.0+).

### Installation

First, run the following code to install the `remotes` package and use it to install the custom `PolandRoundTable` package.

```{r, eval = FALSE}
install.packages("remotes")

remotes::install_github("kbodwin/PolandRoundTable")
```

#### Dependencies

The installation process may prompt you to install some *dependencies*, or other packages that are needed for the `PolandRoundTable` package to function. Even if you already have these packages installed, it is probably a good idea to update them to the most recent version by re-installing.

If you are not comfortable navigating these installs, simply run the following to insure that you have all necessary software:

```{r, eval = FALSE}
prt_dep <- c("igraph",
    "dplyr",
    "magrittr",
    "purrr",
    "lubridate",
    "tidyr",
    "shiny",
    "tidygraph",
    "plotly",
    "shinyWidgets",
    "devtools",
    "shinythemes",
    "ggraph",
    "ggiraph",
    "ggstar")

install.packages(prt_dep)
```

### Producing the App

Once you have finished all installations, you may use the app at any time. Simply open an R console and run

```{r, eval = FALSE}
library(PolandRoundTable)
run_network_app()
```

## Citations

If you make use of the raw datasets in your research, please cite:

> Domber, Gregory F., & Bodwin, Kelly Nicole. (2023). Historical institutional affiliations of participants in Poland's 1989 Round Tablew \[Data set\]. Zenodo. <https://doi.org/10.5281/zenodo.7837475>

If you make use of the Shiny app in your work, please cite the datasets above, as well as the app itself:

> Kelly Nicole Bodwin, Nguyen, Kien, & Domber, Gregory F. (2023). kbodwin/PolandRoundTable: Application for visualizing social networks of Polish 1989 Round Table participants. (v1.0.0). Zenodo. <https://doi.org/10.5281/zenodo.7837452>

#### Feedback and Input

We would love to hear from you about the exciting ways you are using our data and software to explore the history of the Polish Round Table talks!

We would also be interested in hearing feature requests or bug reports, if there are ways we can improve the app or data process.

Drop us a line here with any and all thoughts about this work:

### [Usage Survey](https://forms.gle/iSWeEHVNtJiKPTCA6)
